---
title: "French given names per year per department"
author: "Lucas Mello Schnorr, Jean-Marc Vincent"
date: "October, 2022"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# The problem context
The aim of the activity is to develop a methodology to answer a specific question on a given dataset. 

The dataset is the set of Firstname given in France on a large period of time. 
[https://www.insee.fr/fr/statistiques/2540004](https://www.insee.fr/fr/statistiques/fichier/2540004/dpt2021_csv.zip), we choose this dataset because it is sufficiently large, you can't do the analysis by hand, the structure is simple


You need to use the _tidyverse_ for this analysis. Unzip the file _dpt2020_txt.zip_ (to get the **dpt2020.csv**). Read in R with this code. Note that you might need to install the `readr` package with the appropriate command.

```{r}
library(tidyverse)
```

```{r}
library(DoE.wrapper)
```


```{r}
experiment <- lhs.design( type= "maximin" , nruns= 1000, nfactors= 1 , randomize = TRUE,
digits=0 , seed= 401868 , factor.names=list( Size=c(100,1000000)))
```

```{r}
experiment <- do.call("rbind", replicate(3, experiment, simplify = FALSE))
experiment
```
```{r}
rownames(experiment) <- NULL
experiment
```

```{r}
experiment$Type = "Parallel"
experiment[1001:2000,]$Type = "Sequential"
experiment[2001:3000,]$Type = "Built-in"
experiment
```

```{r}
experiment$order =  sample(as.numeric(rownames(experiment)))
experiment
```
```{r}
ggplot() +               
  geom_point(data = experiment, aes(x=Size, y=order, color = Type))
```
```{r}
experiment <- experiment %>% arrange(order)
experiment
```
```{r}
drops <- c("order")
experiment <- experiment[ , !(names(experiment) %in% drops)]
experiment
```


```{r}
experiment %>% group_by(Type) %>% count()
```


```{r}
write.csv(experiment,"~/workspace/M2R-ParallelQuicksort/data/experiments/experiment.csv", row.names = FALSE)
```


```{r}
file_path = "~/workspace/M2R-ParallelQuicksort/data/experiments/experiment.csv"

experiment_f <- read.csv(file_path)

experiment_f
```

