---
title: "French given names per year per department"
author: "Lucas Mello Schnorr, Jean-Marc Vincent"
date: "October, 2022"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# The problem context
The aim of the activity is to develop a methodology to answer a specific question on a given dataset. 

The dataset is the set of Firstname given in France on a large period of time. 
[https://www.insee.fr/fr/statistiques/2540004](https://www.insee.fr/fr/statistiques/fichier/2540004/dpt2021_csv.zip), we choose this dataset because it is sufficiently large, you can't do the analysis by hand, the structure is simple


You need to use the _tidyverse_ for this analysis. Unzip the file _dpt2020_txt.zip_ (to get the **dpt2020.csv**). Read in R with this code. Note that you might need to install the `readr` package with the appropriate command.

```{r}
library(tidyverse)

```


## Download Raw Data from the website
```{r}
df <- read.csv("data/sama_2014-10-13/measurements_03:47.csv",header=T)
plot(df$Size,df$Time,col=c("red","blue","green")) #[df$Type]
```
```{r}
df
```


```{r}
new_df <- read.csv("data/debian_2022-11-10/measurements_15_21.csv",header=T)
plot(new_df$Size,new_df$Time,col=c("red","blue","green"))

```


```{r}
new_df
```
```{r}
ggplot() +               
  geom_point(data = df, aes(x=Size, y=Time, color = Type))
```
```{r}
ggplot() +               
  geom_line(data = df, aes(x=Size, y=Time, color = Type))
```


```{r}
dplyr::glimpse(new_df)

```

```{r}
ggplot() +               
  geom_point(data = new_df, aes(x=Size, y=Time, color = Type))
```

```{r}
ggplot() +               
  geom_line(data = new_df, aes(x=Size, y=Time, color = Type))
```

